\documentclass[12pt]{report}
\usepackage{geometry}
\usepackage{titling}

\geometry{margin=1in}

\newcommand{\subtitle}[1]{
	\posttitle{
		\par\end{center}
		\begin{center}\large#1\end{center}
		\vskip0.5em}
}

\title{Application Domain Study}
\subtitle{CS 383 - Homework 1 - Group 3}

\author{
Mason Fabel \\
\and Morgan Holbart \\
\and Ronald Rodriguez \\
\and Tessa Saul \\
\and Lance Wells \\
\and Zachary Yama \\
}

\date{\today}

\begin{document}

\maketitle

\chapter{Related Applications}

\begin{section}{Nethack - Mason Fabel}
For the application domain study, I chose to study the classic roguelike
game Nethack. I chose this title for a number of reasons. First of all,
the roguelike genre is, if not the direct predecesor, the inspiration for
the majority of tile-based dungeon crawling games, among others. This being
so, I chose Nethack as it a mature and heavily-played roguelike.

For my study, I am going to play a session of Nethack and record a
condensed version of my adventures. I will then discuss the elements that
I think can be used successfully in our project, followed by the elements
that I think can not.

\begin{subsection}{Gameplay}
Nethack opens with a pretty straightforward selection of menus asking the
player for their class, race, and gender. I selected a human male Tourist,
as I am merely making a small tour of this game for this study. A short,
three paragraph intro is dropped on me, telling me to descend the
dungeon and fetch some amulet.

And that's where the handholding stops. I'm dropped into the game with the
oh-so-helpful message "Be careful!", and nothing else. Some quick fiddling
will reveal that the various keypresses correspond to different actions,
and that pressing '?' will bring up some very cryptic help. Fortunately,
this is far from my first time in Nethack, so I'm over the gigantic
learning curve needed to even make sense of the screen and perform basic
actions.

Thus, off I go, looking for the amulet. The game itself was pretty simple
to watch. Explore the map, look for the stairs to go deeper, run away from
monsters until my pet kills them and I loot the corpses. In this manner I
cleared a few levels and then called it a day.
\end{subsection}

\begin{subsection}{The Good}
There are a lot of good, fun elements to Nethack, even if I didn't come
across too many of them in the short period I had to play and study it.
Some of these features are a good match for the project we are making,
while others are not so good.

The first element of note which could be put to good use in our project
is the fact that each dungeon level is randomly generated, and so are
the items, to a certain extent. What Nethack does is layout a general
level generation scheme (rectangular rooms connected by corridors during
the early levels), and then randomly create floors which follow those
rules. Thus, the game isn't at the complete mercy of the random generator,
but can set and follow certain expectations while still providing variety
and a unique experience each game.

Furthermore, the item's are partially randomized. While each game of
Nethack will have the same potion of acid, ring of levitation, or other
items, but each game the effect of, say, the red potion is different.
Play once, the red potion is healing. Play again, it's hallucination, or
perhaps grease. While this particular mechanic hinges on the fact that a
major part of a game of Nethack is discovering what your items are and
what they do, the general principle of using controlled randomness to add
variety to different playthroughs of the game is praiseworthy, and should,
in my opinion, be used throughout our final project.
\end{subsection}

\begin{subsection}{The Bad}
While Nethack is a fantastic game, some of it's features would translate
extremely poorly over to our project. After all, Nethack is a single-player
dungeon crawler with an emphasis on mechanical depth, and our project
is a cooperative multiplayer game, which calls for what can be very
different features.

The most obvious feature to discard would the be the interface. While
Nethack functions quite well once you grow accustomed to it, the controls
are very obtuse and extremely difficult to learn. To further complicate
the issue, most of the information regarding what is happening and all of
the information regarding what actions the player can take are completely
hidden, unless one already knows a complicated maze of menus and actions.
When we build our project, we should make it very clear what is currently
happening and what actions the player can respond with.

Another feature of Nethack that should definitely be dropped from our
project is permanent player death. In Nethack, if you die then your game is
over, permanently, and you must start again. While that contributes in a
large way to the charm of the game, this is completely out of place in a
multiplayer game, where a single mistake could force the player out of an
otherwise pleasant game. Furthermore, when playing with other people it
is very often fun to compare progress and try to one-up your friends.
Having the game threaten to wipe that out forever is typically not a good
feature for such games.
\end{subsection}

\end{section}

\begin{section}{Teleglitch - Ronald Rodriguez}
For my application domain study, I chose to study Teleglitch. Teleglitch 
is a rogue-like top-down shooter/slasher action game. I chose this game 
as I felt it was somewhat similar to our goal of a tile-based rpg without 
``waiting'' battles.

I will complete the first few levels of Teleglitch and record my thoughts 
about which features I feel would and would not work for our 
game. 

\begin{subsection}{Gameplay}
From the start menu of Teleglitch, the player can choose to play, read a 
tutorial, or choose a level they have previously completed. Once the 
user selects play/continue, they are dropped into the world. Players can
move  the character around the screen using standard keyboard/mouse
controls. 

The way levels are designed and generated in Teleglitch is interesting. Not
only are levels randomly generated each time the user plays, but there is a
neat mechanic where most of the level remains hidden until your character 
actually sees that section by navigating over to it. This causes great 
moments of surprise when you travel to a new part of the level only to be 
met by a horde of enemies. 

The core of the gameplay comes in the form of killing enemies using various
weapons. The character begins the game with a sword that he can use to
melee  enemies with. Other weapons like guns and bombs are obtained via
loot drops from downed enemies and item boxes. One of the main mechanics of
the game comes in the form of being able to combine certain loot items into
more useful weapons. For instance, you may find some nails in an item box
at one point, and then a bit later find a baseball bat after you kill an
enemy. The game allows you to simply combine these two items into a
nail-bat which gives it a damage boost over a standard baseball bat. There
are multitudes of these potential item combinations. 

The game also incorporates elements of the survival horror genre in the
form of ammo being somewhat of a rarity, encouraging the player to think
very carefully about when they choose to use their guns insted of their
melee weapons.
\end{subsection}

\begin{subsection}{The Good}
The shooting mechanics in Teleglitch are exceptional. When a gun is fired
at an enemy, a noticable ``blur'' effect happens to the character and the
bullet itself. This effect adds a palpable ``weight'' to the shots, making
them visceral, impactful, and highly satisfying. 

The loot combining element is an excellent mechanic. It causes players to
think critically about the potential ways they can combine the multitude of
items in their inventory. A great sense of satisfaction comes when you
think of a way to combine two things and the game allows you to do so
exactly as you envisioned. 

The level designs are very cool, having a very industrial, apocalyptic
aesthetic. The levels are somewhat minimal in design, which lends well to
the game overall, as there is rarely any unnecessary clutter obstructing
you from attacking your foes or getting around the environment. 

While it has been a complaint of many industry reviewers, I think the
low-fi look of the game overall is quite charming. A game like this doesn't
need to have super high-res graphics. I felt it would add almost nothing to
the game at all. 
\end{subsection}

\begin{subsection}{The Bad}
The application of ``ammo rarity'' in this game only ended up frustrating
me, rather than giving me a feeling of tension. With Teleglitch, where the
shooting is easily the highlight of the entire game, implementing a
mechanic that lessens the amount of times you will use that awesome gunplay
seems silly to me. 

The melee combat in Teleglitch is nowhere near as satisfying as the
gunplay. There is a noticable lack of ``feedback'' when the melee weapons
connect with enemies, sometimes leaving you unsure if you actually hit the
enemy at all, since most take more than one hit to kill.  

In Teleglitch, when you want to see the whole map of the level you are
currently on, you need to press a button that actually zooms out from the
current room you're in and shows you the whole map. I felt this feature
broke the immersion of the game. I think an onscreen mini-map that 
could be toggled on and off would have been a much better choice both for
Teleglitch and for our game.
\end{subsection}

\end{section}

\begin{section}{Binding of Isaac - Lance Wells}

For my section regarding the application domain study, I have decided to
review general concepts regarding our game's genre, in addition to drawing
from specific examples introduced in popular ``Roguelights''.

To begin, our general game concept runs closely in the vein of a
traditional ``Roguelike'', that is, our game borrows several features from
the genre introduced by the game Rogue. Included in our description
specifically are:

\begin{itemize}
\item Tile-Based Aesthetic and Generation
\item Dungeon Crawling Gameplay
\item Role-Playing Game thematics
\end{itemize}

To facilitate the description of these items, the following is a list of
interpretations of these features from the Roguelight ``Binding of Isaac''.

\begin{itemize}
\item \emph{Tile-Based Aesthetic and Generation}

Binding of Isaac (or more recently, the Rebirth edition), features a
tile-based dungeon layout. The game limits the range of tile-values to only
a select few (Passable Terrain, Impassible Terrain, Destructible Terrain,
Spikes), while altering tilesets throughout each themed floor set. This
strategy effectively simplifies the gameplay while still allowing for a
large variety of room layouts.

\item	\emph{Dungeon Crawling Gameplay}

Rather than adopting a standard free-movement-by-floor interpretation of a
dungeon crawler. Binding of Isaac chooses to instead only display one
``room'' at a time to the player. This effectively lowers the complexity in
displaying an entire ``floor'' simultaneously while still maintaining a
maze-like floor-by-floor dungeon crawler.

\item	\emph{Role-Playing Game thematics}

Binding of Isaac chooses to abandon conventional leveling systems while
instead functioning on a skill-by-skill basis. Each character is influenced
by a series of simple numeric-based "skills" that change the gameplay
slightly (e.g. movement speed, projectile speed, luck, etc.).
Simultaneously, a player may alter their experience by picking up various
stat-altering or otherwise gameplay-altering items ranging from mild to
extreme in effect. What more, Binding of Isaac maintains a small portion of
``risky'' decisions often incorporated into roguelikes similar to
random-effect pills, varying-effect  cards, and health-depleting ``Deals
with the Devil''.
\end{itemize}
\end{section}

\begin{section}{Crawl - Morgan Holbart}
I chose to conduct my domain study on the recently released beta version
of the game ``Crawl''. It has some roguelike elements, but more closely
resembles that of a dungeon crawler. The game is a multiplayer dungeon
crawler where 4 players compete to reach level 10 and defeat the boss
before the other players. 

\begin{subsection}{Gameplay}
The game has, as expected for its genre, very simple controls and graphics,
the value of the game comes from its unique competitive nature and ever
changing strategy. The twist to this dungeon crawler is that one player
is a human, and the other 3 players play a the monsters in the dungeon. To
start each player chooses a god to worship (which gives a passive bonus),
and the game begins. One player is randomly selected and sent off to
explore the dungeon.

Once the player moves from room to room, the other players are given the
ability to control monsters in the room and begin attacking the human
player. They are rewarded blood points for damage dealt to the player, and
the player is granted experience for killing monsters. The goal of the
human is to try to get as much experience as possible to reach level 10
and try to beat the boss, with careful consideration of the fact that the
higher level he gets, when you move on to the next dungeon floor, every
other player will get points to level up their monsters for every level
the human gets. Once the human is killed, whoever dealt the killing blow
will take place as human starting at level 1 (or the previous level he
achieved) and can begin moving on.

All the blood dealt as a monster can be turned to money as a human, and
used to buy gear if you find the store that is on each level. The game
balances itself out by making your monster stronger for every level
another player gains, and by only being able to buy gear by earning money
while playing as a monster.
\end{subsection}

\begin{subsection}{Good Ideas}
The game is fast paced, there is limited setup and vast replayability. It
is focused more on the competitive nature and goal of winning than it is
the roguelike or RPG elements. If every player knows what they are doing,
the games can end anywhere between 15 and 30 minutes with plenty of
push-pull for first place.

The game balances itself, for if you are a human for a long duration the
monsters get enhanced, and if you are the monster for a long duration you
will be enhanced until you surpass the other monsters and can more easily
become the human.

The game is procedurally generated and unpredictable, you never know what
the room you are about to enter will have, it can have an assortment of
traps (controllable by monsters also), monsters, and treasure.

The game has a fairly strong AI, that in its hardest difficulty is fairly
difficult to beat, and plays the game pretty similarly to a human, give or
take a couple bumps in the behavior.
\end{subsection}

\begin{subsection}{Bad Ideas}
The game has some balance issues most likely because of the beta state of
the game. Some monsters are fairly  overpowered compared to others, and
some of the gear can be overpowered, as well. Though careful selection of
monster ugprades can help to counter certain weapons, you can only upgrade
if you have points and at the end of each floor, so the unbalance can
carry the human until he moves on to the next floor.

The boss can be unbalanced based on the weapons and abilities available to
the humans (which are randomly generated). A ranged weapon trivializes the
boss, and a melee weapon is basically impossible to win with against any
competent players because of how the boss is designed.

Level 1 monsters are too weak, and whoever gets human first basically gets
free reign over the first dungeon floor until people can upgrade the
monsters.
\end{subsection}

\begin{subsection}{What Will Translate Well}
The game is a floor based dungeon crawler that is both cooperative and
competitive. Players are working together as monsters, but against the
human, but because of the turn based nature of having control of the human,
it can be difficult to form alliances, thus retaining the competitive
balance. 

The game did procedural generation very well, though sometimes it can be
unbalanced if the current human lands health potions multiple rooms in a
row and another does not.

The game ends fairly quickly, which may or may not be good depending on how
in depth we want the RPG elements to be, if we plan on a fairly low depth
RPG system, the game has replayability and speedy gameplay down.

Unlockables, based on the length of the game, you gain experience which
unlocks new items, gods, and monsters for the game, this adds to the
replayability, and allows you to add strategic depth over time, keeping the
game simple to start and increasing the strategy as you play more and more.
This also gives you a reason to play again (try a new item).
\end{subsection}
\end{section}

\begin{section}{Wanderlust: Rebirth - Tessa Saul}
I chose to look at the tile-based RPG Wanderlust: Rebirth. I chose this
game because of the games I already had, this one best fit the discription
of what we are trying to do. This game has a single-player mode and a co-op
mode with up to four people. 

I played through the tutorial and first couple levels in single-player
mode. 

\begin{subsection}{The Good}
This game has a simple and quick class selection system. There are four
classes and two genders. Each combination has a detailed portrait. Besides
that, the only thing you have to pick to make a new character is a name. 
This makes starting a new co-op game very fast.  
Leveling up was also very easy and fast. The game gives you points to
distribute among your skills. Depending on how you place them, you could 
probably have very different builds. The game has a `lobby' where you can
reassign all of the points that you have earned into different builds, and
then try them out. Another thing I liked about the game is that when I died
multiple times, my character was stunned for a few seconds, and then I got
to continue playing. I didn't even realize that I died until the end of
the level when it deducted points for the death. This means that even if
you are very bad at the game, it will not interrupt the fun you are having
to point out how bad you are. The game fills in your abilities by giving
you AI companions as well. This is nice because you do not have to play a
class to have its abilities, like healing. 
\end{subsection}

\begin{subsection}{The Bad}
The inventory and crafting systems in this game feel very clunky. The
tutorial had me create an item from a recipe, but I had to select all the
items for the recipe by hand in my inventory. This was frustrating because
the items only had little icon type images to represent them, so I had to
look at each item to see if it was one I needed for the recipe. If we have
an inventory system, I think we should look at other games for ideas. 

The controls on the game feel `spammy'. By this, I mean that you can pretty
much press random buttons rapidly and get about the same result as if you
planned out an attack. This is emphasized by the fact that if you play
single-player, the game gives you three AI companions that are the other
classes. They swarm around you when you hold still, and attack everything
at the same time as you, so it is hard to see what you are doing. 
\end{subsection}
\end{section}

\begin{section}{Quest of Dungeons - Zachary Yama}
For my application domain study I chose to analyze the dungeon crawler,
Quest of Dungeons. I chose this title because it’s a more recent, indie
approach to the genre of dungeon crawlers and we will be approaching our
game from the same sort of perspective. It also happens to be fairly
popular, so it would be interesting to see what features make this game a
successful dungeon crawler.
 
The basis of this study will be drawn from a single 15 minute campaign
session. Throughout the session I will record any important game features
and interesting experiences I find. Once a reasonable list has been
compiled I will break it into three sections: how the gameplay directs 
the player, what elements could be included in our game, and elements that
probably shouldn’t be included in our game.
 
\begin{subsection}{Gameplay and Direction}
On game launch, the player is presented with a generic and straightforward
navigation menu. Once a campaign is started, the player is asked to pick a
class. There are four choices: wizard, warrior, assassin, and shaman. After
a class is selected, a short humorous cut scene is presented where the
other three classes that were not selected suggest you take on the world’s
most evil villain and plunge into his dungeon alone. Reluctantly the class
that was chosen by the player exits the scene and the actual game begins.
In my case, it was the Wizard.
 
After being dropped into the game, I noticed a generic set of GUI elements:
health, mana, a minimap, a place to set skills for quick use, and a set of
buttons to open the inventory, stats, and quest menus. After playing with
the keyboard for a bit all the movement controls become clear. The game
uses WASD movement and a point and click system to perform actions.
 
Now that I had an idea of how the mechanics worked I decided to go see if I
could kill something. In fact I could; with giant fireballs. I also noticed
there were many breakable objects lying around. Many yielded gold and even
items and spells I could equip. After running around for a while I found a
large stone that began a quest to kill so and so with some neatly included 
directions. After spending about five minutes heading off on an interesting
adventure I decided to be done and quit the game.
\end{subsection}
 
\begin{subsection}{Useful Features}
Many of the features I experienced in Quest of Dungeons were really fun and
interesting. Though, with such a short session my first impressions may not
do all elements justice. Regardless, there are some elements that would
blend well in our game, and others not so much.
 
The most interesting thing I came across was the spell system. There aren’t
any skills, you have to find them from drops. What’s particularly
interesting is that you can use spells that seem more like something a
warrior would use. The catch is that the spell does damage scaled off of a
specific stat. For example, the fireball skill I had scaled off of
intellect by a rating of B, where the rest of the stats it scaled off of
were F.
 
Another game element I found inspirational was the fast paced combat
system. Monsters do only move when you do, but the way the game behaved was
very polished and quick. This is somewhat attributed to the WASD movement
and the snappy response of having an active spell that is used only by
pointing and clicking. I didn’t need to select anything or think about what
to do; I just did it.

Lastly, the questing system was somewhat interesting. Being able to go out
adventuring and randomly find a stone that gives you newfound direction was
rather refreshing. It would be nice to include some sort of questing system
in our game so that the player isn’t suck with just running around the
dungeon waiting for the last level to come.  
\end{subsection}
 
\begin{subsection}{Harmful Features}
This game was extremely fun to play, but my experiences were with a single
player campaign. There are some components that clearly wouldn’t work for
our multiplayer version, which is what we intend to do.
 
One of these features is locking a player in a room. For a multiplayer
system this would be more harmful than beneficial. If a party accidently
overextends into a higher level room, they would likely die and 
lose valuable experience and have to restart at no real fault of their own.
There would also be inconvenient communication requirements to make sure
that the whole party is ready to move on, and which room they decide as a
whole to move to next. These two elements will needlessly slow down the
players.
 
Another feature that shouldn’t be included is permanent death. If a player
dies, they are required to start completely over. This would prevent
players from wanting to play with each other simply because they don’t
know how prepared the other player really is. This again slows party
progress and requires additional unnecessary communication.
\end{subsection}
\end{section}

\end{document}
